More straightforward movement variable definitions

### 2024-04-03 scripts/01.2_categorizeMovement.qmd

```{r}
pca_movement <- prcomp(movement[,-c(1:2)])
fviz_pca_biplot(pca_movement, axes = c(1,2), 
                geom = c("point"), 
                alpha.ind = 0.3,
                repel = T)+
  theme_classic()
fviz_screeplot(pca_movement, addLabels = T) # that's pretty strongly loaded on the first principal component, so I think I can justify just using that as an overall measure of movement. To double check, let's see the contributions:
contrib_movement <- round(pca_movement$rotation[,1:3]*-100, 2) %>%
  as.data.frame() 
contrib_movement # Okay, PC1 does seem like it can be a general measure of movement.
movement <- movement %>%
  mutate(movement_pc1 = pca_movement$x[,1]*-1)
```

@@

```{r}
pca_roost <- prcomp(roost_behavior[,-c(1:2)])
fviz_pca_biplot(pca_roost, axes = c(1,2), 
                geom = c("point"), 
                alpha.ind = 0.3,
                repel = T)+
  theme_classic()
fviz_screeplot(pca_roost, addLabels = T) # This is also strongly loaded on one principal component, so I feel quite comfortable taking it as a measure of roost diversification/routine
contrib_roost <- round(pca_roost$rotation[,1:3]*-100, 2) %>%
  as.data.frame()
contrib_roost
roost_behavior <- roost_behavior %>%
  mutate(roost_pc1 = pca_roost$x[,1]*-1)
```

@@

```{r}
pca_space <- prcomp(space_use[,-c(1:3)]) # only including home range and coreAreaFidelity (because it would be redundant to include all three)
fviz_pca_biplot(pca_space, axes = c(1,2), 
                geom = c("point"), 
                alpha.ind = 0.3,
                repel = T)+
  theme_classic()
fviz_screeplot(pca_space, addLabels = T) # This is also strongly loaded on one principal component, so I feel quite comfortable taking it as a measure of roost diversification/routine
contrib_space <- round(pca_roost$rotation[,1:3]*-100, 2) %>%
  as.data.frame()
contrib_space
space_use <- space_use %>%
  mutate(space_pc1 = pca_space$x[,1]*-1)
```

>>>>

```{r}
pca_movement <- prcomp(movement[,-c(1:2)])
fviz_pca_biplot(pca_movement, axes = c(1,2), 
                geom = c("point"), 
                alpha.ind = 0.3,
                repel = T)+
  theme_classic()
fviz_screeplot(pca_movement, addLabels = T) # that's pretty strongly loaded on the first principal component, so I think I can justify just using that as an overall measure of movement. To double check, let's see the contributions:
contrib_movement <- round(pca_movement$rotation[,1:3]*-100, 2) %>%
  as.data.frame() 
contrib_movement # Okay, PC1 does seem like it can be a general measure of movement.
movement <- movement %>%
  mutate(movement_pc1 = pca_movement$x[,1]*-1)
save(pca_movement, file = here("data/calcMovement/pca_movement.Rda"))
```

@@

```{r}
pca_roost <- prcomp(roost_behavior[,-c(1:2)])
fviz_pca_biplot(pca_roost, axes = c(1,2), 
                geom = c("point"), 
                alpha.ind = 0.3,
                repel = T)+
  theme_classic()
fviz_screeplot(pca_roost, addLabels = T) # This is also strongly loaded on one principal component, so I feel quite comfortable taking it as a measure of roost diversification/routine
contrib_roost <- round(pca_roost$rotation[,1:3]*-100, 2) %>%
  as.data.frame()
contrib_roost
roost_behavior <- roost_behavior %>%
  mutate(roost_pc1 = pca_roost$x[,1]*-1)
save(pca_roost, file = here("data/calcMovement/pca_roost.Rda"))
```

@@

```{r}
pca_space <- prcomp(space_use[,-c(1:3)]) # only including home range and coreAreaFidelity (because it would be redundant to include all three)
fviz_pca_biplot(pca_space, axes = c(1,2), 
                geom = c("point"), 
                alpha.ind = 0.3,
                repel = T)+
  theme_classic()
fviz_screeplot(pca_space, addLabels = T) # This is also strongly loaded on one principal component, so I feel quite comfortable taking it as a measure of roost diversification/routine
contrib_space <- round(pca_space$rotation[,1:3]*-100, 2) %>%
  as.data.frame()
contrib_space
# Multiply by -1 to flip these around, because right now `pca_space` shows these both being negatively loaded.
space_use <- space_use %>%
  mutate(space_pc1 = pca_space$x[,1]*-1)
save(pca_space, file = here("data/calcMovement/pca_space.Rda"))

# Plots of the extremes
high <- space_use %>% filter(space_pc1 == max(space_pc1, na.rm = T))
low <- space_use %>% filter(space_pc1 == min(space_pc1, na.rm = T))

```
